\documentclass[11pt]{article}
\input{../shared/project_style.tex}
\begin{document}

\docheader{Module 2: Boundary-Driven Compression Physics}{How inward boundary forcing launches waves, compression fronts, and central stagnation}{Purpose: explain the qualitative plasma dynamics produced when the outer boundaries of a magnetized two-dimensional domain are driven inward in time.}

\begin{overviewbox}
\textbf{What this note does.} This note gives the physical story that connects the imposed boundary drive to the interior compressed state. The central ideas are finite-speed communication, compressive-wave propagation, front interaction, stagnation, and the loss of uniformity when the drive is mistimed or overly violent.
\end{overviewbox}

\symboltableheader
$\mathbf{n}$ & outward unit normal & Unit vector normal to a boundary face.\\
$v_n$ & normal velocity & Fluid velocity component normal to a boundary.\\
$f_t(t)$ & temporal pulse profile & Time-dependent drive envelope.\\
$f_s(s)$ & spatial profile & Drive variation along a boundary coordinate $s$.\\
$c_s$ & sound speed & Thermal compressive-wave speed.\\
$v_A$ & Alfv\'en speed & Magnetic-tension wave speed.\\
$c_f$ & fast magnetosonic speed & Characteristic speed of the fastest compressive MHD disturbance.\\
$T_{\text{pulse}}$ & pulse duration & Characteristic time over which the boundary drive is active.\\
$\tau_r$ & rise time & Characteristic turn-on time of the pulse.\\
$\Omega_{\text{core}}$ & core region & Central region used to judge compression quality.\\
\symboltableend

\section{Physical picture of inward boundary forcing}
Suppose one wall is commanded to impose an inward normal velocity. The plasma adjacent to that wall cannot respond everywhere at once. Instead, the boundary perturbs only the nearby fluid first, and that perturbation is then communicated inward through MHD wave propagation.

In a purely hydrodynamic gas this would be described as a compressive pressure disturbance. In ideal MHD the response is richer because gas pressure and magnetic stresses both participate. The interior therefore learns about the drive through a magnetized compressive front rather than through instantaneous action at a distance.

\begin{physicsbox}
\textbf{Core idea.} The boundary does not squeeze the entire domain directly. It launches inward-propagating MHD disturbances, and the compressed state forms only after those disturbances travel, interact, and redistribute mass, momentum, pressure, and magnetic flux.
\end{physicsbox}

\section{Weak compression waves versus strong fronts}
If the imposed drive is smooth in time and modest in amplitude, the boundary typically launches a relatively smooth compressive wave. The plasma variables increase gradually across the front, and the solution remains easier to interpret physically and numerically.

If the drive is turned on too abruptly, or if the imposed inward speed becomes too large relative to the characteristic signal speeds, the front steepens. In that regime the flow becomes more shock-like: gradients sharpen, reflections become stronger, and the final compressed state often becomes less uniform.

For an interview discussion, the practical distinction is simple. A gentle drive is useful for understanding the underlying compression logic. A harsh impulsive drive may create stronger peak compression, but it also introduces more structure, more residual motion, and more sensitivity to numerics.

\section{Why signal speed matters}
The relevant signal speed is not arbitrary. In a magnetized plasma, the fastest compressive information is often associated with the fast magnetosonic speed $c_f$, which combines gas-pressure and magnetic effects.

That speed matters in at least three ways:
\begin{enumerate}
    \item it sets how quickly a front launched at the edge can reach the core,
    \item it controls whether a chosen pulse is better thought of as a slow squeeze or as an impulse, and
    \item it sets the natural scale for comparing amplitude and timing across different runs.
\end{enumerate}

If the pulse duration is short compared with the inward travel time, the plasma experiences a more impulsive hit. If it is comparable to or longer than the transit time, the core experiences a more sustained squeeze. Those two situations can lead to very different interior states even when the net injected momentum is similar.

\section{Interaction of fronts launched from multiple sides}
The main project scenario is not a single-sided push. It is a multi-sided inward compression problem. If all four walls are driven inward symmetrically, then four inward-traveling compression fronts move toward the center.

In the clean symmetric picture:
\begin{itemize}
    \item each front has the same strength and timing,
    \item the fronts converge toward the geometric center,
    \item the core densifies and magnetizes without a preferred direction, and
    \item the strongest residual velocity should diminish near the center after interaction.
\end{itemize}

This interaction event is the heart of the repository. The project is not only about launching fronts. It is about what kind of interior compressed region those fronts build when they meet.

\section{Central stagnation and compressed-core formation}
When inward-directed flows collide near the center, the flow cannot continue converging indefinitely. The interior speed must drop, and momentum must be redirected into pressure buildup, magnetic-field amplification, or secondary motions.

That slowing of the inward flow is the core of \emph{stagnation}. In a good compression state, the repository should not show only a sharp density spike. It should show a compressed central region with relatively low residual core velocity. Otherwise the result is more like transient sloshing than useful compression.

This is why stagnation metrics belong next to compression metrics. A run that reaches a large peak $\rho$ but leaves strong organized flow in the core may be scientifically less attractive than a run with slightly smaller peak compression but cleaner stagnation.

\section{Why asymmetry spoils uniform compression}
Perfectly symmetric compression is fragile. Several small asymmetries can bias the interior state:
\begin{itemize}
    \item one wall may start earlier than another,
    \item one wall may be driven with larger amplitude,
    \item the spatial drive profile may be stronger near the corners than near the middle,
    \item the background magnetic field may make compression respond differently parallel and perpendicular to the field.
\end{itemize}

When that happens, the fronts no longer meet under equivalent conditions. The nominal stagnation point shifts, the compressed region becomes skewed, and the interior can retain a directional bias.

\begin{keybox}
\textbf{Important physical message.} Uniform compression is not determined only by how much total inward impulse is applied. It also depends on how that impulse is distributed in time, in space, and among the different walls.
\end{keybox}

\section{Magnetic anisotropy during compression}
Because the system is magnetized, compression is not purely isotropic even if the domain and the drive are symmetric. Magnetic pressure resists local pileup wherever $|\mathbf{B}|$ grows, while magnetic tension redistributes stress along field lines.

As a result, the orientation of the initial magnetic field matters. Compression perpendicular to the field tends to crowd magnetic flux more directly, whereas compression along the field can couple differently through tension and pressure. This does not invalidate the symmetric baseline; it simply means that the field direction is itself part of the physics.

\section{Why smooth pulse shaping is physically useful}
The temporal pulse shape is not a cosmetic detail. It determines how abruptly momentum is injected from the boundary.

A discontinuous turn-on can launch artificial high-frequency transients that say more about the forcing prescription than about the intended compression experiment. A smooth pulse reduces that problem. It also makes it easier to interpret the relation between the drive and the resulting front structure.

The physically important pulse-shape knobs are:
\begin{itemize}
    \item rise time,
    \item duration near peak forcing,
    \item fall time,
    \item repetition interval if a pulse train is later introduced.
\end{itemize}

Those knobs change whether the domain experiences one broad compressive event, one sharp impact, or several interacting inward pushes.

\section{Expected qualitative evolution of the baseline case}
For the first symmetric four-sided inward pulse, the expected evolution is:
\begin{enumerate}
    \item the boundary forcing turns on smoothly,
    \item inward-traveling compression fronts launch from all four walls,
    \item density, pressure, and magnetic-field strength rise progressively toward the interior,
    \item the fronts interact near the center,
    \item a compressed core forms, and
    \item diagnostics decide whether the core is strong, uniform, symmetric, and near stagnation.
\end{enumerate}

That sequence is the reference picture against which later drive asymmetry, pulse shortening, pulse trains, or stronger forcing should be compared.

\section{Concluding summary}
Boundary-driven compression is fundamentally a finite-speed communication problem in magnetized fluid dynamics. The boundaries act first, the interior responds later, and the final quality of the compressed core depends on wave speed, front strength, symmetry, pulse shape, and magnetic anisotropy. That is why the project treats boundary forcing as a design variable rather than as a mere boundary-condition detail.

\end{document}
