\documentclass[11pt]{article}
\input{shared/project_style.tex}
\lhead{MHD\_BoundaryCompression}
\rhead{Physics Notes Outline}
\begin{document}

\DocTitle{Physics Notes Outline}{Interview-ready physics note set for the boundary-compression project}{Purpose: define a compact textbook-style note family that explains the governing equations, compression physics, drive families, magnetic-flux amplification, diagnostic metrics, and baseline reference case in a physics-first format.}

\begin{overviewbox}
\textbf{Document purpose.} This outline is no longer only a placeholder list. It is the narrative plan for the note set that can feed an interview slide deck directly. The intended progression is: governing equations $\rightarrow$ compression physics $\rightarrow$ drive taxonomy $\rightarrow$ magnetic-field amplification $\rightarrow$ quantitative metrics $\rightarrow$ baseline case.
\end{overviewbox}

\section{Symbol and acronym table}
\begin{symtable}
MHD & magnetohydrodynamics & Fluid model that couples plasma motion to magnetic-field evolution. \\
EOS & equation of state & Closure relation connecting pressure, energy, and density. \\
$\rho$ & mass density & Primary compression variable for the plasma fluid. \\
$p$ & thermal pressure & Scalar gas pressure that rises during compressive heating. \\
$\mathbf{v}$ & fluid velocity & Bulk flow field produced by the inward drive and later stagnation. \\
$\mathbf{B}$ & magnetic field & Field whose amplification and uniformity are part of the scientific objective. \\
$\Phi_B$ & magnetic flux & Magnetic flux through a chosen surface or flux tube cross section. \\
$v_A$ & Alfv\'en speed & Characteristic MHD wave speed associated with magnetic tension. \\
$c_f$ & fast magnetosonic speed & Relevant signal speed for compressive disturbances. \\
$v_n$ & normal boundary velocity & Imposed inward speed in the first drive model. \\
$J$ & objective function & Composite scalar used to rank compression quality. \\
\end{symtable}

\section{Purpose of the physics-note set}
This note family is meant to function as part of the scientific project itself, not as optional packaging. Each module should behave like a short textbook chapter: define notation, explain the physical picture, state the core equations, and connect the physics to the code architecture and diagnostics.

\section{Module 1: ideal-MHD compression problem}
\subsection{Scientific purpose}
Define the governing equations and explain what boundary-driven compression means when the computational grid itself is fixed. This note answers the basic question: how can a plasma become compressed on a fixed Eulerian mesh if the boundary does not literally move the grid points?

\subsection{Topics that belong in this note}
\begin{itemize}[leftmargin=1.8em]
    \item meaning of ideal MHD as a conducting-fluid model,
    \item distinction between primitive and conservative variables,
    \item conservative-form governing equations,
    \item ideal-gas equation of state,
    \item magnetic pressure and magnetic tension during compression,
    \item frozen-in magnetic-flux intuition,
    \item why density and magnetic-field magnitude can intensify together.
\end{itemize}

\section{Module 2: boundary-driven compression physics}
\subsection{Scientific purpose}
Explain the dynamical sequence launched by an inward boundary drive. The note should distinguish between smooth compressive waves and stronger shock-like fronts, and it should show how inward-propagating fronts interact when launched from multiple sides.

\subsection{Topics that belong in this note}
\begin{itemize}[leftmargin=1.8em]
    \item inward forcing picture,
    \item finite-speed communication through the plasma,
    \item weak-wave versus strong-front logic,
    \item role of signal speeds in setting interior response,
    \item central stagnation and compressed-core formation,
    \item how asymmetry in timing or strength creates nonuniform compression.
\end{itemize}

\section{Module 3: boundary-drive models}
\subsection{Scientific purpose}
Define the idea of a drive family and explain why an abstraction layer is needed. The reader should come away understanding the difference between a physical drive prescription and a numerical boundary implementation.

\subsection{Topics that belong in this note}
\begin{itemize}[leftmargin=1.8em]
    \item prescribed inward velocity as the first model,
    \item pressure loading as a future model,
    \item moving-wall boundaries as a future model,
    \item pulse trains, timing offsets, and spatial tapering,
    \item mapping from drive intent to ghost-cell states.
\end{itemize}

\section{Module 4: uniformity metrics and objective functions}
\subsection{Scientific purpose}
Define what ``uniform compression'' means mathematically and physically. A key job of the document is to explain why large compression alone is not enough.

\subsection{Topics that belong in this note}
\begin{itemize}[leftmargin=1.8em]
    \item core-region averaging,
    \item density, pressure, magnetic, and total-pressure compression ratios,
    \item coefficient-of-variation style nonuniformity metrics,
    \item symmetry penalties,
    \item residual core-speed and stagnation metrics,
    \item composite objective-function structure,
    \item fair comparison of competing drive prescriptions.
\end{itemize}

\section{Module 5: baseline case definition}
\subsection{Scientific purpose}
Document the first reference problem in full detail. It should be possible for a reader to understand the baseline setup and expected qualitative evolution without reading the MATLAB code.

\subsection{Topics that belong in this note}
\begin{itemize}[leftmargin=1.8em]
    \item square-domain geometry,
    \item initial density, pressure, and magnetic field,
    \item adiabatic index,
    \item which sides are driven and in what inward directions,
    \item temporal pulse shape,
    \item spatial profile along each boundary,
    \item expected compression history and first diagnostic plots to inspect.
\end{itemize}

\section{Module 6: flux compression and field amplification}
\subsection{Scientific purpose}
Provide a dedicated explanation of why inward plasma compression can amplify the magnetic field in ideal MHD. This note should be especially useful for interview presentation because it links the boundary drive directly to the magnetic objective of the project.

\subsection{Topics that belong in this note}
\begin{itemize}[leftmargin=1.8em]
    \item magnetic-flux intuition,
    \item simple flux-tube scaling argument,
    \item perpendicular versus parallel compression,
    \item relation to magnetic pressure and magnetic tension,
    \item limits of the frozen-in picture,
    \item connection back to the repository diagnostics.
\end{itemize}

\section{Suggested derivation notes}
The repository can later contain a derivation subcollection including:
\begin{itemize}[leftmargin=1.8em]
    \item \texttt{conservative\_ideal\_mhd\_equations.tex},
    \item \texttt{boundary\_velocity\_prescription.tex},
    \item \texttt{compression\_ratio\_metrics.tex},
    \item \texttt{magnetic\_flux\_compression\_measures.tex}.
\end{itemize}
These documents would go beyond implementation and walk through the derivation logic step by step.

\section{Suggested implementation notes}
The implementation-note set should complement, not replace, the physics notes. The most useful early implementation documents remain:
\begin{itemize}[leftmargin=1.8em]
    \item run order and workflow,
    \item boundary-drive interface,
    \item ghost-cell strategy,
    \item diagnostics pipeline.
\end{itemize}

\section{Concluding interpretation}
The physics-note set should be treated as an interview-facing scientific asset. In a repository like this one, the value of the project depends not only on the solver but also on whether the governing physics, drive abstractions, and diagnostic definitions are explicit enough that a reader can understand the logic without reverse-engineering the MATLAB code.

\end{document}
